---
title: "4_26_17_homework"
author: "Allie Gaudinier"
date: "4/25/2017"
output: 
  html_document: 
    keep_md: yes
---
### Homework for April 26th, 2017
###Sections 3.1- 3.51 in R for Data Science

Install the tidyverse package once:

install.packages("tidyverse")

```{r}
library(tidyverse)

summary(mpg)

#learn more about this dataset
?mpg

#plot displ by hwy
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))


#general code for basic ggplot functions
#ggplot(data = <DATA>) + 
  #<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

##3.2.4 Exercises
```{r}
ggplot(data = mpg)
#this produced an empty plot

#check number of rows and columns in mtcars
nrow(mtcars)
ncol(mtcars)

?mpg
#drv. f = front-wheel drive, r = rear wheel drive, 4 = 4wd

ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
##this plot shows that there is fwd, rwd, and 4wd for the types of cars
```



####Color by car class
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

#### Warning: Using size for a discrete variable is not advised.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

####alpha aesthetic controls the transparency of the points
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

####shape of the points
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.

make all the points blue
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

Figure 3.1 R has 25 built in shapes that are identified by numbers

##3.3.1 Exercises

```{r}
#1. 
#ggplot(data = mpg) + 
  #geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
#to make all points  = "blue" should be geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

#2.
#Categorical: manufacturer, model, trans, year, cyl, class, fl, drv
#   Continuous: displ, cty, hwy
str(mpg)

#3. #color continuous variable
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))

#4. #map two aes to a variable - it works but looks weird
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, alpha = cty))

#5. what does stroke in aes do? modifies the width of the border
#this only works with some shapes - those that have borders
?geom_point
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, stroke = 1))

#6. makes a graph with true and false colors 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```

###3.5.1 Exercises

```{r}
#1. You can't facet on a continuous variable because it can't be broken into categories or segments

#2. 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))

#this displays the displ by hwy at the in seperate graphs that have been designated by the drv ~ cyl formula
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

#3. The plots are formed by only one category (drv or cyl)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

#4. # It is easier to see the distribution of datapoints in the individual plots, 
#but it is harder to make comparisons
# A larger dataset would make it harder to make comparisons if every point is plotted
#on the came graph
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

#5.
#?facet_wrap
#nrow = number of rows, ncol = # columns - this controls the number of 
#rows or columns in the facet plot
#?facet_grid
#facet_grid doesn't have this because it uses the variables instead

#6. When using facet_grid() you should usually put the variable with more unique levels in
#the columns. Do this because they are easier to compare than variables that have closer
#values - the rows are better for this comparison.
```




