---
title: "5_10_17_homework"
author: "Allie Gaudinier"
date: "5/9/2017"
output: 
  html_document: 
    keep_md: yes
---

##5.2.4 Exercises

1. Find all flights that: 
    1. Had an arrival delay of two or more hours
```{r}
library(nycflights13)
library(tidyverse)
filter(flights, arr_delay >= 120)
```

  2. Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

  3. Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA","DL"))
```

  4. Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% c(7,8,9))
```

  5. Arrived more than two hours late, but didn’t leave late    
```{r}
filter(flights, arr_delay > 120, dep_delay == 0)
```

  6. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

    7. Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
?between()
filter(flights, between(dep_time, 0, 600))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent? Missing - arrival time , etc
```{r}
count(flights, dep_time == NA)
```

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
```{r}
NA ^ 0
NA | TRUE
FALSE & NA
NA * 0
```

Everything ^ 0 is 1
NA or TRUE, therefore it is TRUE
FALSE and NA, can be FALSE
NA*0 is not defined


##5.3.1 Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, desc(is.na(x)))
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_time)
```

3. Sort flights to find the fastest flights.
```{r}
arrange(flights, air_time)
```

4. Which flights travelled the longest? Which travelled the shortest?
```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```

##5.4.1 Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, starts_with("dep"), starts_with("arr"))
# you could also ! select everything besides these - but that's a lot of unnecessary typing
```

2. What happens if you include the name of a variable multiple times in a select() call?
It only prints it once
```{r}
select(flights, dep_time, dep_delay, dep_time)
```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
selects variables based on variables in character vector.
This could be useful if you just want to select a specific year, month, etc
```{r}
?one_of()
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
Case doesn't matter in the select function default.
```{r}
select(flights, contains("TIME", ignore.case = F))
```

##5.5.2 Exercises

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
transmute(flights, dep_time, dep_time_hour = dep_time %/% 100, dep_time_min = dep_time %% 100, dep_time_sincemidnight = dep_time_hour*60 + dep_time_min)
```

2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
air_time is in minutes, arr and dep time is in hours:minutes - need to convert
```{r}
transmute(flights, air_time, altairtime = arr_time - dep_time)

#this should fix it
transmute(flights, air_time, dep_time_sincemidnight = (dep_time %/% 100)*60 + dep_time %% 100, 
          arr_time_sincemidnight = (arr_time %/% 100)*60 + arr_time %% 100)
```

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related? dep_delay + sched_dep_time = dep_time. Time zones, dateline, overnight flights are all a problem
```{r}
transmute(flights, dep_time, sched_dep_time, dep_delay)
```

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank(). ties.method = "min"
```{r}
transmute(flights, dep_delay, carrier, flight, delay_rank = min_rank(desc(dep_delay)))
```

5. What does 1:3 + 1:10 return? Why? 
It cycles through the 1:3 vector for the additions
```{r}
1:3 + 1:10
```

6. What trigonometric functions does R provide?
R computes the cosine, sine, tangent, arc-cosine, arc-sine, arc-tangent, and the two-argument arc-tangent


