---
title: "5_17_17_homework"
author: "Allie Gaudinier"
date: "5/16/2017"
output: 
  html_document: 
    keep_md: yes
---


```{r}
library(nycflights13)
library(tidyverse)

delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
delays
```

##Exercises 5.6.7


1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:
 A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>%
  group_by(flight) %>%
  summarise(early = mean(arr_delay <= -15), late = mean(arr_delay >= 15)) %>%
  filter(early ==0.5 , late == 0.5)
```
 
 A flight is always 10 minutes late.
```{r}
 filter(not_cancelled, arr_delay == 10)
```

 A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
```{r}
not_cancelled %>%
  group_by(flight) %>%
  summarise(early = mean(arr_delay <= -30), late = mean(arr_delay >= 30)) %>%
  filter(early ==0.5 , late == 0.5)

```
 99% of the time a flight is on time. 1% of the time it’s 2 hours late.
```{r}
not_cancelled %>%
  group_by(origin, carrier, flight) %>%
  summarise(quantile(arr_delay > 120, 0.99))
```
 
Which is more important: arrival delay or departure delay?
Arrival delay - can make up time in the air. Look at depature delay, arrival delay - not always equal

2. Come up with another approach that will give you the same output as 

  
```{r}
not_cancelled %>% count(dest)

##same as above
not_cancelled %>%
group_by(dest) %>%
      summarise(flights = n())
```

```{r}
not_cancelled %>% count(tailnum, wt = distance)

##same as above
not_cancelled %>%
  group_by(tailnum) %>%
  summarise(flights = sum(distance))
```

3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?
Using two variables is a little sloppy - The best way to show if the flight existed is air_time

4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
cancelled <- flights %>% 
  filter(is.na(dep_delay), is.na(arr_delay))

library(ggplot2)
cancelled %>%
  group_by(day) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = T))
 
#ggplot(, mapping = aes(x = flight, y = avg_delay)) + 
#    geom_point(alpha = 1/10)
```

5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

```{r}
not_cancelled %>%
  group_by(carrier) %>%
  summarise(mean_arr_delay = mean(arr_delay, na.rm = T)) %>%
  arrange(desc(mean_arr_delay))
```


6. What does the sort argument to count() do. When might you use it?
?count()
count(x, ..., wt = NULL, sort = FALSE) 
Sorts the results ofthe count from highest to lowest

##5.7.1 Exercises

1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.
These function within the groups

2. Which plane (tailnum) has the worst on-time record?
```{r}
not_cancelled %>%
  group_by(tailnum) %>%
  summarise(avg_arr_delay = mean(arr_delay), na.rm = T) %>%
  arrange(desc(avg_arr_delay))
```

3. What time of day should you fly if you want to avoid delays as much as possible? - May 7th
```{r}
not_cancelled %>%
  group_by(hour) %>%
  summarise(avg_arr_delay = mean(arr_delay), na.rm = T) %>%
  arrange((avg_arr_delay))

```

4. For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination.
```{r}
not_cancelled %>%
  group_by(dest) %>%
  summarize(total = sum(arr_delay))

##look at stacey's code for the second answer
```


5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.
```{r}
not_cancelled %>%
  group_by(origin, year, month, day) %>%
  mutate(prec_flight_delay = lag(dep_delay))

## look at min-yaos code
```



6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

```{r}
#look at staceys answer
not_cancelled %>%
  group_by(dest) %>%
  mutate(fastest = min(air_time)) %>%
  ungroup() %>%
  arrange(air_time)
```


7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.
```{r}
flights %>%
  group_by(dest) %>%
  count(carrier) %>%
  summarise(num_carrier = n ()) %>%
  filter(num_carrier >= 2)
```


8. For each plane, count the number of flights before the first delay of greater than 1 hour.
not_cancelled %>%
  group_by(flight)
  summarize()
  Look at Michelle's answer for this
