---
title: "5_31_17_homework"
author: "Allie Gaudinier"
date: "5/31/2017"
output: 
  html_document: 
    keep_md: yes
---

##10.5 Exercises

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame). onlt the first 10 rows would print if it was a tibble.
print(mtcars)

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration? Depending on how much data you ask for a dataframe will give you a vector (one column) or a df (mutliple columns)

```{r}
library(tidyverse)

df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
df[,1:2]
```

```{r}
df2 <- tibble(abc = 1, xyz = "a")
df2$x
df2[, "xyz"]
df2[, c("abc", "xyz")]
```


3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

mpg[var]
mpg[[var]]
mpg[,var]
get(var, mpg)
select(mpg, matches(var))
subset(mpg, select= var)


4. Practice referring to non-syntactic names in the following data frame by:
  Extracting the variable called 1.
  
```{r}
testt <- tibble(
  `a` = c(1,5,7,9,11,15,18),
  `b` = c(3,5,8,3,12,16,7)
)

testt[["a"]]
```


Plotting a scatterplot of 1 vs 2.
```{r}
library(ggplot2)
ggplot(testt, aes(x = `a`, y = `b` )) +
 geom_point()
```


Creating a new column called 3 which is 2 divided by 1.

```{r}
testt[["c"]] <- testt[["a"]] + testt[["b"]]
```


Renaming the columns to one, two and three.
```{r}
testt <- rename(testt, one = `a`, two = `b`, three = `c`)
glimpse(testt)
```


5. What does tibble::enframe() do? When might you use it?
?enframe()
A helper function that converts named atomic vectors or lists to two-column data frames. For unnamed vectors, the natural sequence is used as name column.

```{r}
enframe(c(x =1, y =2, z = 12))
```

6. What option controls how many additional column names are printed at the footer of a tibble?
?print.tbl_df
n_extra	: Number of extra columns to print abbreviated information for, if the width is too small for the entire tibble. If NULL, the default, will print information about at most tibble.max_extra_cols extra columns.