---
title: "Untitled"
author: "Allie Gaudinier"
date: "5/2/2017"
output: 
  html_document: 
    keep_md: yes
---

## May 3rd Homework

###Exercises 3.6.1

```
1.  
    geom for line chart: geom_line()
    geom for boxplot: geom_boxplot()
    geom for histogram: geom_bar()
    geom for area chart: geom_area()
```
```{r}
#2. This code will plot the points for all the cars data and create a trend line for each 
#type of drive
library(ggplot2)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

#3. show.legend = FALSE will take the legend away from the plot
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(se = FALSE, show.legend = FALSE)

#4. se - display confidence interval around smooth? (TRUE by default, see level to control
#?geom_smooth()

#5. These plot are the same - the aes are specified in the ggplot function or the geom functions
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

#6. 
#plot 1
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(se = FALSE, show.legend = FALSE)

#plot 2
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth((mapping = aes(group = drv, show.legend = FALSE)), se = FALSE)


#plot 3
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth((mapping = aes(group = drv)), se = FALSE)

#plot 4
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(show.legend = FALSE, se = FALSE)

#plot 5
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(mapping = aes(group = drv), se = FALSE, show.legend = TRUE)

#plot 6
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(size = 3, shape = 21, stroke = 3, color = "white", mapping = aes(fill = drv))
```

###3.7.1 Exercises

```{r}
#1. #?stat_summary
#This produces a boxplot type plot with distributions that can be marked
ggplot(data = mpg, aes(drv, cyl)) + 
  stat_summary()
```
```
2. What does geom_col() do? How is it different to geom_bar()?
geom_col() is short-hand for geom_bar(stat = "identity")
geom_bar() is for bar plots

3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?
#on this website :http://sape.inf.usi.ch/quick-reference/ggplot2/geom


4. What variables does stat_smooth() compute? What parameters control its behaviour?
computes a trendline using lm or glm

5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?
These graphs show the proportion of diamonds per category but not proportion per total number of diamonds per category
```
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = 1))
```

###3.8.1 Exercises
```
1. What is the problem with this plot? How could you improve it?
There isn't much data - label the types of cars. Also, not showing all the points
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = class)) + 
  geom_point() +
  geom_jitter()
```

```
2. What parameters to geom_jitter() control the amount of jittering?
It adds a small amount of random variation to the location of each point, and is a useful way of handling overplotting caused by discreteness in smaller datasets. Width and height: 
Amount of vertical and horizontal jitter. The jitter is added in both positive and negative directions, so the total spread is twice the value specified here.

?geom_jitter()

3. Compare and contrast geom_jitter() with geom_count().
geom_count increases the size of a datapoint if there are multiple counts for that exact point
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = class)) + 
  geom_point() +
  geom_jitter()
  
  ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = class)) + 
  geom_point() +
  geom_count()
```

```
4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.The default position is stacked
```
```{r}
ggplot(data = mpg, mapping = aes(x = class, color = drv)) + 
  geom_bar()
```

###3.9.1 Exercises
```
1.Turn a stacked bar chart into a pie chart using coord_polar().
```
```{r}
ggplot(data = mpg, mapping = aes(x = class, fill = drv)) + 
  geom_bar() +
  coord_flip()
```

```
2. What does labs() do? Read the documentation. 
?labs()
Change axis labels and legend titles

3. What’s the difference between coord_quickmap() and coord_map()?
?coord_quickmap() -  a quick approximation that does preserve straight lines. It works best for smaller areas closer to the equator
?coord_map() - projects a portion of the earth, which is approximately spherical, onto a flat 2D plane using any projection

4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?'
?geom_abline() - These paired geoms and stats add straight lines to a plot, either horizontal, vertical or specified by slope and intercept
?coord_fixed() - A fixed scale coordinate system forces a specified ratio between the physical representation of data units on the axes. The ratio represents the number of units on the y-axis equivalent to one unit on the x-axis.
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

###4.4 Practice
```
1.Why does this code not work?
Spelling error

2. Errors fixed:
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl = 8)
filter(diamonds, carat > 3)

3. Alt + shift + k opens references for keyboard shortcuts
```

### Plot my own data

```
Load data for root system architecture measuremens
```
```{r}
Set1 = read.table("~/Desktop/RootPhenotyping/Final_Combined_Analysis/TDNA_40a.txt", header=T)
Roots1 = data.frame(Set1, stringsAsFactors=TRUE )
head(Roots1)
```

```
Make boxplots for the data at 1mM and 10mM nitrate compared to Col-0
```
```{r}
#PR Length
ggplot (data = Roots1, mapping = aes( x = NO3_Level, y = logPR, fill = Genotype)) +
  #geom_violin(position = position_dodge(width=0.9), alpha = 0.5) +
  geom_boxplot(position = position_dodge(width=0.9)) +
  geom_point(position=position_jitterdodge(dodge.width=0.9))

#LR number
  ggplot (data = Roots1, mapping = aes( x = NO3_Level, y = logLR, fill = Genotype)) +
  #geom_violin(position = position_dodge(width=0.9), alpha = 0.5) +
  geom_boxplot(position = position_dodge(width=0.9)) +
  geom_point(position=position_jitterdodge(dodge.width=0.9))
  #geom_count(position=position_jitterdodge(dodge.width=0.9))
  
  #Total LRL
  ggplot (data = Roots1, mapping = aes( x = NO3_Level, y = logLRL, fill = Genotype)) +
  #geom_violin(position = position_dodge(width=0.9), alpha = 0.5) +
  geom_boxplot(position = position_dodge(width=0.9)) +
  geom_point(position=position_jitterdodge(dodge.width=0.9))
  #geom_count(position=position_jitterdodge(dodge.width=0.9))